---
title: "Desicion tree assignment"
author: "Andrii Taborovskyi"
date: "16 05 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

This assignment is based on materials from [mlcourse.ai] (https://mlcourse.ai)

## Building a decision tree for predicting heart diseases

Let's read the data on heart diseases. The dataset can be downloaded from the course repo from [here](https://www.dropbox.com/s/kr1d9vsnax5wxfo/mlbootcamp5_train.csv?dl=0).

**Problem**

Predict presence or absence of cardiovascular disease (CVD) using the patient examination results.

**Data description**

There are 3 types of input features:

- *Objective*: factual information;
- *Examination*: results of medical examination;
- *Subjective*: information given by the patient.

| Feature | Variable Type | Variable      | Value Type |
|---------|--------------|---------------|------------|
| Age | Objective Feature | age | int (days) |
| Height | Objective Feature | height | int (cm) |
| Weight | Objective Feature | weight | float (kg) |
| Gender | Objective Feature | gender | categorical code |
| Systolic blood pressure | Examination Feature | ap_hi | int |
| Diastolic blood pressure | Examination Feature | ap_lo | int |
| Cholesterol | Examination Feature | cholesterol | 1: normal, 2: above normal, 3: well above normal |
| Glucose | Examination Feature | gluc | 1: normal, 2: above normal, 3: well above normal |
| Smoking | Subjective Feature | smoke | binary |
| Alcohol intake | Subjective Feature | alco | binary |
| Physical activity | Subjective Feature | active | binary |
| Presence or absence of cardiovascular disease | Target Variable | cardio | binary |

All of the dataset values were collected at the moment of medical examination.

## Task

Read the data
```{r}
mlbootcamp <- read.csv("mlbootcamp5_train.csv", sep=";")
```

Transform the features: 
- create "age in years" (`age_years`)dividing age by 365.25 and taking floor ($\lfloor{x}\rfloor$ is the largest integer that is less than or equal to $x$)
- remove "age" feature
- transfrom `cholesterol` and `gluc` to factor

```{r}
mlbootcamp$age_years <- floor(mlbootcamp$age/365.25)
mlbootcamp <- subset(mlbootcamp, select=-c(age, id))
mlbootcamp$cholesterol <- as.factor(mlbootcamp$cholesterol)
mlbootcamp$gluc <- as.factor(mlbootcamp$gluc)
mlbootcamp$cardio <- as.factor(mlbootcamp$cardio)

is.factor(mlbootcamp$cholesterol)
is.factor(mlbootcamp$gluc)
```

Split data into train and test parts in the proportion of 7/3.
The target feature is `cardio`

```{r}
#install.packages('caret')
library(caret)

set.seed(2344)
trainIndex <- createDataPartition(mlbootcamp$cardio, p = 0.7, list = FALSE, times = 1)
train <- mlbootcamp[trainIndex, ]
test  <- mlbootcamp[-trainIndex, ]

head(train, 3)
head(test, 3)
```

Train a decision tree on the dataset `training` with **max depth equal to 3**. For setting this parameter use function `rpart::rpart.control`. Use default cross calidation parameters.

```{r}
#install.packages('rpart')
#install.packages('e1071', dependencies=TRUE)
library(rpart)
library(caret)
train1 <- train(cardio ~ ., data = train, method = "rpart", control = rpart.control(maxdepth = 3))

train1
```

Plot final tree with `rattle` library

```{r}
#install.packages('rattle')
library(rattle)
plot1 <- fancyRpartPlot(train1$finalModel, main="cardio")

plot1
```

What is the accuracy of prediction on test data?

```{r}
prediction1 <- predict(train1, test)
accuracy <- sum(prediction1 == test$cardio) / nrow(test)

accuracy
```
Now set cross validation parameters: `method = "cv"`, `number = 5`. Tree parameters set to default.

```{r}
controlOfATrain <- trainControl(method = "cv", number = 5)
train2 <- train(cardio ~ ., data = train, method = "rpart", trControl = controlOfATrain)

train2
```

Plot final tree with `rattle` library

```{r}
library(rattle)
plot2 <- fancyRpartPlot(train2$finalModel, main="cardio")

plot2
```

What is the accuracy of prediction on test data?

```{r}
prediction2 <- predict(train2, test)
accuracy2 <- sum(prediction2 == test$cardio) / nrow(test)

accuracy2
```
Does the accuracy became better or not?

**No, It remained the same**